### 2015-05-08 ms.Rmd
Setup paper to build via remake

---
title: "TREE: A package for modelling plant TRait Ecology and Evolution"
author: "Daniel Falster, Rich FitzJohn, Mark Westoby"
affiliation: Biological sciences, Macquarie University NSW 2109
output: pdf_document
csl: downloads/style.csl
bibliography: refs.bib
---

# Abstract

Population dynamics in forest communities are strongly size-structured: larger
plants shade smaller plants while also expending proportionately more energy
on building and maintaining woody stems. Although the importance of size-
structure for demography is widely recognised, many mechanistic models either
omit it entirely, or include only coarse approximations. In this paper we
outline an extensible framework and package (called TREE, for TRait Ecology
and Evolution) for modelling plant demography, driven by a mechanistic model
of plant function, across the entire life cycle. At its core, TREE is an
individual-based model, wherein plants are modelled as a system of coupled
differential equations. Individuals from multiple species can be grown in
isolation, in patches of competing plants, or in  metapopulations under a
prevailing disturbance regime. The dynamics within patches of competing plants
are resolved using either 1) a stochastic simulator, or 2) a deterministic
approximation of their density. The latter is achieved via a novel extension
of the Escalator Boxcar Train technique. The combined effects of trait-, size-
and patch-structured dynamics can then be integrated into a population level
estimate of reproductive fitness. The core of TREE is programmed in C++ with
wrappers allowing both interactive and scripted use from R. Moreover, we allow
for alternative physiologies and hyper-parametrisation on the basis of plant
functional traits. A detailed test suite is provided to ensure accuracy. We
discuss how the methods outlined can be used to address fundamental questions
on how functional traits influence the growth of individual plants, whole patches
and assembly of ecological communities.

# Introduction

Problem is that only one dominant vegetation model includes size structure, even
though competition for light widely acknowledged as most fundamental of processes.
Methods for studying size-structured metapopulations poorly developed, relying mostly
on simulation.
Means that unable to answer many questions, particular about traits, where need to
evaluate affect across entire life cycle, and taking into account effects on growth,
fecundity and mortality.

What does tree model do

- Individual plants growing
  - estimate plant growth rates in relation to size, light env, traits
- Plants competing in a patch (stochastic, deterministic)
  -  multi-species self-thinning, successional transitions
- Trait, Size and Patch structured vegetation
  - emergent behaviours: vegetation (e.g. Falster et al 2011), demography
- Fitness in metapopulation
  - community assembly, evolution, adaptation
- Evolution, community assembly, e.g. successional diversity

Series of nested uses

- has multiple possible spin-offs, we highlight some ways of extended
- Each step down is just one way of extending thing above

# The methods

# The approach

Briefly about design.

- Extension of @Falster-2011. - separate
- Major new piece of software (lines), designed with..... 1300 R code, 7000 cpp
- New method for solving PDEs
  + estimate density using modified version of EBT, integration along characteristics.
  + not sure if better than original EBT but simpler
  + at present don't have cohort mean implemented, but could easily enough [RF ~
  2 days] --> would
  need this is wanted to make any claims about one method being better
- stochastic and deterministic versions
- multiple models physiological models possible
- hyper-parametrisation with respect to traits
- Possible to drive core from C++ (with minimal modification), and thereby interfacing other languages

Differences to EBT

- density via point estimates rather than averages across cohorts
- In original EBT, total number of individuals constrained through formation of a
cohort,, but only via first-order approximation. Here estimates of n is exact,
but could be error in total number of individuals.
- adaptive spacing of cohorts

# Use cases

Examples of biological problems solved using the software / How would other people use the model (evolutionary simulations, demographic, etc)

- consider a small code example, expanded on in the vignette

## Demographics

Example from growth trajectories

## Scaling from trait to patch and vegetation

Successional dynamics

Emergent mortality, growth, size distributions

## Community assembly

- Fundamental versus realised niche
- Niche differentiation and evolved neutrality
- Evolutionary responses to environmental change

# Closing comments

Lists of unexplored research questions

- questions about complementarity

List of possible extensions

- other resources
- physiology

\newpage

# Figures

\begin{figure}[ht]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{output/empty.pdf}
\caption{\textbf{Schematic.} Similar to Figure 1 of Falster 2011.
\label{fig:schematic}}
\end{figure}

\newpage

\begin{figure}[ht]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{output/empty.pdf}
\caption{\textbf{Growth trajectories.} something biological to point at, possibly stochastic and deterministic overlaid.
\label{fig:trajectories}}
\end{figure}

\newpage

# Supplementary material

# References