### 2015-08-24 ms.tex
Minor comments from reviewers

\documentclass[a4paper,11pt]{article}
\usepackage[osf]{mathpazo}
\usepackage{ms}
\usepackage[]{natbib}
\raggedright

\newcommand{\plant}{\texttt{plant}}
\newcommand{\smurl}[1]{{\footnotesize\url{#1}}}
\usepackage{graphicx}

\title{{\plant}: A package for modelling forest trait ecology \& evolution}
\author{Daniel S. Falster\textdagger\textasteriskcentered, Richard G. FitzJohn\textdagger, and Mark Westoby}
\affiliation{
Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia\\
\textdagger These authors contributed equally.\\ 
\textasteriskcentered Email for correspondence: \texttt{daniel.falster@mq.edu.au}\\
A manuscript in consideration as a research paper for
publication in MEE as part of the Special Feature \emph{Demography
  beyond the Population}.\\
Word count: ~3900 words}
\date{}

\bibliographystyle{mee}

\usepackage[title,titletoc,toc]{appendix}

\mstype{Applications Note}
\runninghead{The {\plant} package}
\keywords{demography, emergent, fitness, growth,
physiology, metapopulation, mortality, reproduction, size-structure,
tradeoff}

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Population dynamics in forests are strongly size-structured:
  larger plants shade smaller plants while also expending
  proportionately more energy on building and maintaining woody stems.
  Although the importance of size-
  structure for demography is widely recognised, many mechanistic models
  either omit it entirely, or include only coarse approximations.
\item
  Here we introduce the {\plant} package, an
  extensible framework for modelling size- and trait-based demography, 
  ecology and evolution in simulated forests.
  At its core, {\plant} is an
  individual-based mechanistic model where plant physiology and demography are mediated by
  traits. Individuals from multiple species can be grown in isolation,
  in patches of competing plants, or in metapopulations under a
  disturbance regime. Dynamics within patches of competing plants are
  resolved using novel extensions of the ``Escalator Boxcar Train''
  technique. Combined effects of trait-, size- and patch-structured
  dynamics are integrated into metapopulation level estimates of
  reproductive fitness and vegetation structure.
\item
  {\plant} is an open source R package and is available at
  \href{https://github.com/traitecoevo/plant}{github.com/traitecoevo/plant}.
  While accessed from R, the core routines in {\plant} are written in C++.
  The package provides for alternative physiologies and for
  hyper-parameterisation to capture trade-offs among parameters. A
  detailed test suite is provided to ensure correct behaviour of the code.
\item
  {\plant} provides a transparent platform for investigating how  
  physiological rules and functional trade-offs interact with competition and 
  disturbance regimes to influence vegetation demography, structure and 
  diversity.
\end{enumerate}

\section{Introduction}\label{introduction}

Plant growth and demography are fundamentally size- and 
trait-structured, influencing dynamics over time-scales ranging from
instantaneous physiological effects to long-term evolutionary outcomes
\citep{Harper-1977, Westoby-2002}. 
%
As an individual increases its leaf area, it increases its potential
to generate photosynthate.
%
In opposition to this, as individuals grow they must use increasing
fractions of their photosynthetic income to building tissues other
than new leaves; for example, to maintaining support tissues
\citep{Givnish-1988, Enquist-2007} or to reproduction
\citep{Thomas-2011}.
%
Consequently rates of growth, mortality and reproduction change with
individual size \citep{Muller-2006, Ruger-2011, Thomas-2011}.  
Ontogenetic 
patterns of growth also vary with traits; for example,
leaf and wood construction costs influence growth rates 
\citep{Wright-2010}, while seed size and height
at maturation determine the start and end points of ontogenetic
trajectories \citep{Westoby-2002}.
%
Strong feedbacks emerge between individuals within a forest via
competition for light and other resources, such that the growth rate
of one individual depends on the size and traits of nearby
individuals \citep{Shugart-1980, Pacala-1996}.
%
Such feedbacks make it difficult to link observable species traits to
phenomena such as self-thinning, successional transitions, trait
evolution and species coexistence, without
modelling both individual growth and the competitive interactions between
individuals.

Although the importance of size-structure for vegetation dynamics and
diversity has been long been recognised \citep[e.g.,][]{Harper-1977,
  Shugart-1980, Huston-1987}, current research in this area is
dominated by models and theory that either omit size-structure
entirely, or only include coarse approximations. 
%
Theoretical investigations exploring questions about niche-based
species coexistence and ecological drift (neutrality) both rely
primarily on unstructured population models \citep[e.g.][]{MacArthur-1967,
  Tilman-1985, Geritz-1995, Hubbell-2001, Calcagno-2006}.
%
It is common to assume competitive outcomes are influenced by
size-related traits such as adult and seed size, but detailed
size-structure is rarely considered, at least in plant models
\cite[for animal examples, see][]{Deroos-1988, Deroos-1992}.
% TODO: Andrea notes that this section does not flow well.  I agree,
% but let's leave it until after submission.
Similarly, most large-scale models of vegetation dynamics do not
include size-structured demography \citep{Cramer-2001, Sitch-2003,
  Purves-2008,Dekauwe-2014}.  During the 1980s, size-structured
``gap'' models were widely used
\citep[e.g.][]{Shugart-1980,Huston-1987}. However, the next generation
of models focused more heavily on biogeochemistry at the expense of
demography \citep[e.g.][]{Woodward-1995,Haxeltine-1996,Cramer-2001, 
Sitch-2003}. While the
importance of size-structure is starting again to be recognised
\citep{Moorcroft-2001, Purves-2008, Falster-2011, Dekauwe-2014}, all but a couple
of the most popular,
large-scale vegetation models lack any real demography or
size-structure \citep[see][]{Moorcroft-2001, Smith-2014}.

In this note, we introduce the {\plant} package for R \citep{R-2015};
a mechanistic framework for studying the effects of size structure and
trait variation on the demography of individual plants, of patches of
competing plants and of meta-populations structured by a prevailing
disturbance regime.ƒ
%
We developed {\plant} primarily to further investigate if and how species
differing in traits are able to coexist with one another
\citep[following][]{Falster-2011,Falster-2015}.  Size-structured models
provide unique opportunities for studying coexistence via differentiation in
successional strategy \citep{Moorcroft-2001,Huston-1987}.

At the same time, we expect the {\plant} model to be useful for
modelling other demographic phenomena. The internal structure of the {\plant}
model is similar in approach to the ``Ecosystem Demography''
\citep{Moorcroft-2001} and \textsc{lpj-guess} \citep{Smith-2014}
models. Like {\plant}, these models include size-structured
competition within a metapopulation of ``patches''. However, while
these models rely on simulations and were designed to capture large-scale regional
biogeochemistry driven by climatic variables, we designed
{\plant} so that the rules driving the dynamics of the
system could be manipulated and emergent demographic behaviours
studied deterministically and in detail. In {\plant}, it is possible to grow plants in
isolation, in patches of competing plants and in disturbance-driven
metapopulations. The package is thus designed to bridge the gap between
a large-scale coupled simulation model and a simple
abstract theoretical device.

Below we describe the general approach of the package, then provide
sections with details at nested levels of ecological organisation.
Detailed technical documents are provided as supplementary information
(see Appendices); updated versions of these technical documents are
available within the {\plant} package itself.

\section{The approach}

{\plant} is an individual-based model, meaning the dynamics of the population arise from rules specifying how individuals grow and interact. The
core rules in {\plant} are about the short-term physiological
functioning of an individual plant and how these are influenced by its
traits, size and light environment. By being driven by traits, the
model can be extended to potentially very many species. Users will
also be able to add their own physiological models, though this is a
work in progress.

On top of this physiological model, the package implements methods for
population and adaptive dynamics (Fig.  \ref{fig:schematic}) following methods
described by \citet{Falster-2011} and \citet{Falster-2015}. Demographic
phenomena can be studied at the three levels: individual plants, stands of
competing plants and entire metapopulations. The dynamics at these higher
levels of organisation arise as emergent properties, driven by growth
physiology, competition for light and disturbance (Fig.  \ref{fig:schematic}).
We model the emergent behaviour of patches deterministically, based on a
continual  size-density distribution. As such, specifying a patch size
is not required.

\section{Effects of size, trait and environment on demography of individual plants}

% NOTE: Add something saying: we define a species as a group of
% individuals with identical traits?
The core of {\plant} is a model for an individual species'
physiological strategy as specified by its traits
(Fig. \ref{fig:schematic}a). This model estimates rate of biomass
production for a plant, given its size and the current light
environment. Gross photosynthetic income is calculated from the total
leaf area and the light distribution across the plant's canopy. Costs
of tissue respiration and turnover are then subtracted. The remaining
biomass is then allocated between growth and reproduction.  The key
outputs needed by higher levels of the model (see below) are height
growth rate, mortality rate and rate of seed production.  However
additional quantities are also computed, such as total assimilation
and the respiration, turnover and allocation rates for different
tissues (see Appendix \ref{sec:plant})

The default physiological model used in {\plant} largely reflects that
presented in \citet{Falster-2011,Falster-2015}, but with extensions
allowing for diameter growth also to be estimated. We refer to
this default model as \texttt{FFW16}, reflecting our initials in this
paper.  To model
trait variation properly, different parameters (such as rates of leaf turnover,
respiration, photosynthesis per leaf area) in the model are linked via
assumed trade-offs (see Appendix \ref{sec:parameters}).  For example, the trait
\textsc{lma} (leaf mass per unit leaf area) is used to estimate the
rate of leaf turnover, based on an observed scaling relationship
spanning across diverse vegetation and plant functional types
\citep{Wright-2004} (see Appendix \ref{sec:FFW16} for details).

The current parametrisation of the physiological model is based on
\citet{Falster-2011}. Parameter values were drawn from available
literature (see Table 3 in Appendix \ref{sec:FFW16}); however data for
some parameters were limited. Providing updated parameter sets to the
algorithm is
trivial (see Appendix \ref{sec:parameters}); the challenge is to infer
suitable values by constraining outputs from {\plant} against diverse
data sources \citep{Lebauer-2012, Keenan-2013}. This is an area in
which we are still actively working and one of the major use cases we
envisage for {\plant}.

While we have implemented one physiological model (\texttt{FFW16}),
the package is designed to allow arbitrary additional physiologies to
be introduced.  That is, rather than varying the parameters of the
equations, the equations themselves can be changed and the higher
level functionality of the package can still be used.  As an
intermediate level, we allow for changes in key parameters to trigger
changes in other parameters (i.e. hyper-parametrisation), allowing
straightforward modelling of trade-offs.

\subsection{Applications}

With the basic physiological model, {\plant} can be used to estimate
essential physiological rates for individual plants
(Fig. \ref{fig:plant}). The function \texttt{grow\_plant\_to\_size}
takes a given strategy and light environment and grows the plant,
producing a trajectory of size over time (Fig. \ref{fig:plant}a). This
is achieved by integrating an ordinary differential equation of
size-dependent growth rate (see Appendices \ref{sec:demography} and
\ref{sec:plant} for more details). As the plant grows, allocation to
different tissue types varies and the composition of the plant
changes to include more stem and less leaf
(Fig. \ref{fig:plant}b). This in turn affects maintenance and turnover
costs and the growth rate of plants (Fig. \ref{fig:plant}c).  This
shift in allocation alone generates a distinctive hump-shaped pattern
of absolute growth with size \citep{King-2011} and equally the widely
observed decline in relative growth rate with size from birth onwards
\citep{Enquist-2007}.

By varying the light environment and measuring growth rate, the whole
plant light compensation point (\textsc{wplcp}) can be computed (Fig.
\ref{fig:plant}d). The \textsc{wplcp} is the light level at which the plant 
stops growing and is increasingly regarded as the most useful
measure of a species's shade tolerance
\citep{Givnish-1988, Baltzer-2007, Lusk-2013}. As expected, \textsc{wplcp}
increases with plant size, due to increased costs of building and
maintaining stem and leaf tissues \citep{Givnish-1988}. Likewise, \textsc{wplcp}
decreases with \textsc{lma}, because high \textsc{lma} species have lower leaf turnover
\citep{Baltzer-2007, Lusk-2013} (Fig.
\ref{fig:plant}d).

\section{Plants competing in a patch}

Within patches of competing plants, competition for light generates
strong non-linear feedbacks on growth, survival and reproduction. In the
\texttt{FFW16} physiological model, we consider only the effect of
shading on rates of biomass production, although competition for other
resources such as nitrogen or water could also be considered by extending the 
physiological model.

We assume that patches are vertically but not horizontally
structured. In other words, we account for size differences but not spatial 
layout within the patch. Similar assumptions are found in many models
simulating size-structured dynamics
\citep{Shugart-1980,Kohyama-1993, Huston-1987, Moorcroft-2001, Smith-2014}. In addition, 
we assume that patches are very large -- this means the population dynamics 
within the patch behave deterministically \citep{Deroos-1997} and can therefore 
be approximated via a Partial Differential Equation (\textsc{pde}) (see Appendix
\ref{sec:demography}). The same \textsc{pde} captures average behaviour
across a large number of small patches \citep{Moorcroft-2001}. {\plant} may 
therefore capture average demographic behaviour across a range of patch sizes.

Ideally, one would also consider spatial interactions within patches,
however such models are very computationally demanding and also
impossible to solve in a deterministic fashion \citep{Shugart-1980, 
Pacala-1996}. Moreover, it remains unclear whether resolving spatial details within a
patch would provide further useful ecological detail, in part because the
process of competitive thinning tends to break down spatial clusters
\citep{Strigul-2008}.

When modelling a patch, the focus is on changes to the size-
density-distribution \(N(x, H, a)\) with patch age \(a\), i.e. the density of
individuals with traits \(x\) and height \(h\) within a patch of age \(a\).
Density is measured as individuals per unit height and per unit ground area.
Modelling the development of  \(N(x, H,a)\) requires that both the initial size
distribution and the inflow of new recruits be specified. We have primarily
been interested in dynamics of a patch recovering from disturbance and have
therefore started with an empty patch and constant flow of seeds from a global
seed rain (Fig. \ref{fig:schematic}b).

Our approach for modelling size-structured population dynamics is
based on the Escalator Boxcar Train (\textsc{ebt}) technique
\citep{Deroos-1988, Deroos-1992, Deroos-1997}. The \textsc{ebt} solves
the \textsc{pde} describing development of \(N\) by approximating the
density function with a collection of cohorts spanning the size
spectrum. Following a disturbance, a series of cohorts are introduced
into each patch. These cohorts are then transported along the
characteristic equations of the \textsc{pde}. Biologically these are
the growth trajectories of individuals, conditioned on them surviving
(e.g. Fig. \ref{fig:patch}a). Meanwhile, growth and mortality combine
to alter the density of individuals along the trajectories
(Fig. \ref{fig:patch}b).

We introduce two extensions to the \textsc{ebt}: i) a new method for
estimating the size-density distribution and ii) a technique for
handling strong size-asymmetric competitive feedbacks, such as occur
under strong competition for light (see Appendix \ref{sec:demography}
for details). The original \textsc{ebt} proceeds by approximating the
first and second moments of the density distribution \(N\) within each
cohort \citep{Deroos-1997}.
%
In {\plant} we use a different approach. Instead of tracking first and
second moments of the size-distribution within cohorts, we track a
point mass estimate of \(N\) along the trajectory for the boundary of
each cohort. We found this approach preferable because it does not
require us to maintain a separate set of equations for the smallest
cohort.

The extension for handling strong size-asymmetric competition involves
adaptive refining of time points at which new cohorts are introduced
into the population. In the original \textsc{ebt}, cohorts were
introduced at a fixed time interval. Under competition, the growth
trajectories of individuals born at similar times can diverge
substantially over time (Fig. \ref{fig:patch}a; Appendix
\ref{sec:demography}, \ref{sec:cohort-spacing} and
\ref{sec:patch}). This can become a problem later during patch
development because much of the population can be located between
widely-spaced cohorts, causing low numerical accuracy. To maintain
numerical accuracy, we use an iterative algorithm to adaptively refine
the cohort spacing until the trajectories of adjacent cohorts remain
adequately resolved throughout development of the patch (see Appendix
\ref{sec:demography} for details). This refinement causes an uneven
distribution of cohort introduction times across patch age (Fig.
\ref{fig:patch}a), with considerably tighter spacing of cohorts in
younger patches.

\subsection{Applications}

Modelling of size-structured dynamics via solving of a deterministic
\textsc{pde} enables multiple demographic phenomena to be
investigated. First, {\plant} can simulate demographic 
phenomena typically observed in patches
recovering from a disturbance, including self-thinning and
successional replacement. Whereas previous approaches for modelling
self-thinning were limited to a single species
\citep[e.g.][]{Barnes-2004, Coomes-2007}, {\plant} accommodates
multiple species with different traits (Fig. \ref{fig:schematic}b;
Fig.  \ref{fig:patch}). Moreover, these features emerge from the 
combined effects of physiology and competition for resources, rather 
than being prescribed in the model. Second, {\plant} allows for
the combined effects of physiology and succession on emergent 
characteristics of the entire patch to be investigated
\citep{Falster-2011}. Leaf area cover and rates of biomass 
production tend to vary  non-linearly with patch 
age \citep{Binkley-2002, Smith-2001, Ogawa-2010};
in {\plant} such outcomes arise naturally as emergent properties, 
driven by demography (Fig. \ref{fig:patch}c).

\section{Trait, size and patch structured vegetation}

All vegetation is subject to a disturbance regime, such that patches
of the landscape are cleared at some rate, by fire, cyclone od
landslide.  The vegetation thus comprises a series of patches
differing in time since last disturbance and linked via seed dispersal
(Fig.  \ref{fig:schematic}b). Such a system is often referred to as a
structured metapopulation \citep{Gyllenberg-2001}.

In {\plant} and following \citep{Kohyama-1993, Moorcroft-2001,
  Falster-2011} we assume an infinite number of patches, all experiencing
the same disturbance regime and sharing a common seed dispersal pool
(i.e., the island model; Fig. \ref{fig:schematic}b). We also assume
that the disturbances remove all established vegetation.  With this
approach we can scale up from simulating a single patch to simulate an
entire metapopulation.  The frequency-density \(P(a)\) of patches aged
\(a\) in the landscape is needed. Assuming there are a large number of
patches means the dynamics of \(P\) behave deterministically and can
be approximated via a second \textsc{pde} (see Appendix
\ref{sec:demography} for details). Scaling from patch to
metapopulation is then achieved by weighting the temporal dynamics of
a single patch by \(P(a)\), the relative abundance of patches aged
\(a\) in the metapopulation (Fig. \ref{fig:patch}c).

The main numerical challenge in stepping from a single patch to a
metapopulation is to identify the equilibrium seed rain of the
metapopulation. An ``input'' seed rain is required to simulate the
demography of the metapopulation, this in turn produces an ``output''
number of seeds. The equilibrium occurs when, for each species,
a given input seed rain
produces an equivalent output seed rain. Finding this equilibrium
requires finding the root of a multi-dimensional function, subject to
the constraint that this root is stable (e.g., the trivial equilibrium
of zero seed rain always exists, but is often not stable; Appendix
\ref{sec:equilibrium}).

An attractive feature of the metapopulation concept implemented here
is that integrates both equilibrium and non-equilibrium approaches to
modelling ecological dynamics \citep{Kohyama-1993, Moorcroft-2001,
  Falster-2011}. An equilibrium may be approached at the level of
metapopulation, meaning the seed rain and size structure across the
entire metapopulation is approximately stable. Yet the structure of
vegetation within individual patches is constantly in flux. Patches
continue to age and their residents continue to grow until they are
disturbed.

\subsection{Applications}

The dynamic equilibrium captured by the trait-, size- and
patch-structured metapopulation model potentially resolves several
challenges faced by unstructured models. First, by scaling up from
patch to metapopulation, we close the demographic feedback loop to
create a self-regulating system. Unlike some vegetation models, there
is no need to set the density of seeds or seedlings.  These densities
arise as outputs from the model, based on the combined effects of the
model's physiological rules, species traits, competitive interactions
and the prevailing disturbance regime.

% NOTE: Andrea points out that young patches may not be the most
% common patch type in the metapopulation - it will depend on the
% regime.  And on the definition of young.
Second, the metapopulation concept allows for the effects of disturbance
regimes on vegetation structure and dynamics to be properly incorporated. In
{\plant} characteristics for the entire vegetation (metapopulation) -- such
as total leaf area, biomass and productivity -- are obtained by averaging
patch-level outcomes over the frequency-distribution of patch ages. Because
young patches tend to be more common in the landscape, metapopulation averages are
strongly influenced by the structure of younger patches (e.g.
Fig. \ref{fig:patch}c).  Similarly, the distributions of  abundance and growth rate
observed in large forest  plots \citep[e.g.][]{Muller-2006, Coomes-2007} arise
as  emergent phenomena. Moreover, the predictions from {\plant}
are for a frequency distribution of patch states across the landscape
(Fig. \ref{fig:emergent}, see more detail in Appendix \ref{sec:emergent}). As such,
size- and patch-structured models provide unique opportunities for linking with
satellite and forest survey data \citep{Moorcroft-2001, Purves-2008}.

Finally, with the extension to the metapopulation level, we are
modelling demography across the entire life cycle and are thus in a
position to estimate invasion fitness -- the ability of a rare
individual with particular traits to establish in a community of
established residents at their equilibrium densities
\citep{Metz-1992, Geritz-1998}.  While ideally calculated as the
long-term per capita growth rate, a more convenient indicator of
fitness in structured metapopulations is given by the
logarithm of the basic reproduction ratio, \(R\)
\citep{Gyllenberg-2001, Metz-2001}.
The basic reproduction ratio is simply the total seed output of
the new type, averaged across the metapopulation. So a new type can
invade when \(\log R > 0\).

With this measure of fitness, we can model trait-based adaptive
dynamics and community assembly of forest communities, following
 \citep{Falster-2015}. This is best
thought of as occurring on a fitness landscape, a surface with fitness
plotted against one or more species traits
(Fig. \ref{fig:fitness}). The reproductive success of individuals
possessing a particular trait value depends on both the traits and
densities of other species in the current resident community. Fitness
landscapes will be positive in regions that allow invasion, and zero
for species that are at equilibrium. The slope of the fitness
landscape with respect to traits indicates the direction of selection
(dashed line in Fig. \ref{fig:fitness}a) indicates selection for
smaller trait values (see Appendix \ref{sec:fitness} for
details). Fitness landscapes are constantly reshaped as species are
introduced and evolve, reflecting the non-linear demographic feedbacks
that occur under frequency-dependent evolution. Stable coexistence is
possible when species with different traits can mutually invade
\citep{Geritz-1998, Chesson-2000}. Following the fitness gradient may lead to an
evolutionary branching point \citep{Geritz-1998}; at
this point there is no directional selection (solid line in
Fig. \ref{fig:fitness}a) but invasion is possible both above and below
it in trait space. It is often possible to find a configuration where
fitness is negative everywhere (solid line in Fig. \ref{fig:fitness}b)
except for the zero points where species are present. This community
is immune to invasion from any other strategies and represents a
prediction for the trait mixture that might might be favoured by
natural selection \citep{Geritz-1998}.

Identifying conditions promoting trait diversity is a valuable use-case
for {\plant}. The simple example here can be extended to multiple dimensions
and can be used to investigate how trade-offs facilitate coexistence
\citep{Falster-2015}. Co-occurring plant species differ in a range of physiological
traits, yet the conditions allowing for stable coexistence remain largely
unknown \citep{Adler-2013}.

\section{Implementation details}

{\plant} is written in C++ and R, using the package \texttt{Rcpp}
\citep{Eddelbuettel-2011, Eddelbuettel-2013} to bridge between the two
languages. Due to the computationally intensive nature of the model,
the core physiological and demographic components are in C++ to
maximise speed. We used templated types that allow the demographic and
evolutionary components to be driven by alternative physiological
models while retaining the same interface at higher levels. Despite
mostly being written in C++, all components of the model, down to
individual plants, can be extracted and interacted with dynamically in
R, using the \texttt{RcppR6} package \citep{RcppR6}.

{\plant} makes use of much existing software, including the R
computational environment \citep{R-2015}, the R packages \texttt{Rcpp}
\citep{Eddelbuettel-2011, Eddelbuettel-2013}, \texttt{R6}
\citep{Chang-2014}; and the Boost Library for C++
\citep{Schaling-2014} via the R package \texttt{BH}
\citep{Eddelbuettel-2015}. Source code is hosted at
\href{https://github.com/traitecoevo/plant}{github.com/traitecoevo/plant}.
Installation instructions are available at this webpage, binary
releases will also be available there.  (Note that {\plant} does not
currently compile on Windows, but this will change soon as R core
upgrade the Windows C++ compiler).

\section{Closing comments}

The {\plant} package makes it easier for researchers and forest
practitioners alike to investigate a variety of non-linear demographic
phenomena.  Rather than provide a model with a few entry points, we
have developed an extensible framework that we hope can be used to
answer a number of questions at a variety of scales. The specific
use-cases highlighted here arise from our interest in the assembly of
trait mixtures in communities. However a variety of other uses would
be possible to address other aspects of community assembly. For
example, we expect the package will prove useful for i) investigating
how morphological and physiological traits influence growth rate and
shade tolerance, ii) emergent phenomena such as multi-species self-
thinning, successional transitions and age-related changes in
vegetation productivity, and iii) as a bridge between empirical data,
large-scale simulators of global biogeochemical cycles and simple
abstract theoretical models.

\section{Acknowledgements}

DS Falster was supported by an Australian Research Council (ARC)
discovery grant (DP110102086). RG FitzJohn was supported by the
Science and Industry Endowment Fund (RP04-174). M Westoby was
supported by a fellowship from the ARC. We thank {\AA}
Br{\"a}nnstr{\"o}m, J Camac, A De Roos, U Dieckmann, R Duursma, J
Johansson and C Prentice for helpful discussions. The authors have no 
conflicts of interest to declare.

\clearpage
\bibliography{refs}

\clearpage

\section{Figures}\label{figures}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/schematic}

\caption{Processes modelled within {\plant} include physiological, population and
evolutionary dynamics.
Dynamics across all three levels are driven by the
physiological sub-model determining an individual's physiological and
demographic rates on the basis of its traits, size and light environment (a).
(b) Competitive hierarchies  are modelled by tracking the height distribution of
individuals within a ``patch''. The intensity
of shading indicates the density of individuals at a given height for
different species (distinguished by colours). A size- and patch-structured
metapopulation consists of a distribution of patches linked by seed dispersal.
Disturbances occasionally remove all vegetation within a patch, resetting the
patch. (c) The traits of the resident species determine the shading environment
across the metapopulation, which in turn determines the fitness landscape.
Resident traits adjust through directional selection
and through the introduction of new species. Adapted from
\citet{Falster-2011}, \citet{Falster-2015}. }

\label{fig:schematic}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/plant.pdf}
\caption{Physiological dynamics for individual plants varying in size,
  trait and light environment. Solid lines are for the high
  \textsc{lma} (leaf mass per unit area) strategy, dashed lines are
  for the low \textsc{lma} strategy.  (a) Growth trajectories are
  influenced by both light environment and traits.  (b) Over time, the
  fraction of living tissue switches from leaf towards sapwood, with
  high \textsc{lma} species having relatively more mass in leaf than
  low \textsc{lma} species. (c) Size-dependent growth rates
  (derivative of panel (a)) peak at around 5m high, but this position
  varies with both trait and light level.  (d) Declines in growth rate
  with light vary with plant size and traits.  Whole plant light
  compensation points (indicated by dots) emerge as the point where
  growth rate is zero (x intercept).  The light level used in panels
  (a) and (c) is indicated with coloured ticks.  See Appendix
  \ref{sec:plant} for further details and code.}
\label{fig:plant}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/patch.pdf}
\caption{Population dynamics for two species competing within a patch.
  Red and blue lines are low and high \textsc{lma} species from
  Fig. \ref{fig:plant}.  Lines in panel (a) show growth trajectories
  for individuals on the boundary of successive ``cohorts'' for each
  species, using a schedule of introduction times generated by
  {\plant}'s adaptive algorithm.  The initial growth rate advantage of
  the low \textsc{lma} species (see Figure \ref{fig:plant}c) means it
  quickly over tops the high \textsc{lma} species, suppressing its
  growth.  Self-thinning creates space for the high \textsc{lma}
  species to establish below the canopy of the faster species. Panel
  (b) shows the same trajectories as in panel (a), but with shading
  indicating the density $N$ of individuals at given size and patch
  age (light = low density, dark = high density).  (c) The leaf area
  index at ground level for each species.  The dotted line is the
  total leaf area index (summed across both species), and the lines on
  the right side are average when integrated over the whole
  metapopulation structure.  See Appendix \ref{sec:patch} for further
  details and code.}
\label{fig:patch}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/emergent.pdf}
\caption{Emergent size-distributions and demography across a metapopulation.
Grey lines show patterns within individual patches differing in time
since disturbance (age), with blue lines indicating patches of a
specific age. Red lines show average relationship across the metapopulation, 
obtained by weighting the grey lines by patch and individual abundance. 
Panels show (a) patterns of abundance, (b) leaf area and
(c) height growth rate against size. Units of panels (a) and
(b) are amount per unit height per unit ground area.  
See Appendix \ref{sec:emergent} for further details and code.}
\label{fig:emergent}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/fitness.pdf}
\caption{Example fitness landscape for \textsc{lma} (leaf mass per
  unit leaf area), showing potential for stable coexistence of
  multiple types.  Fitness is the log of the population growth rate --
  at zero species are at equilibrium, above zero they will increase in
  density.  (a) Landscape generated by a single species (indicated by
  dots).  At \textsc{lma} of 0.1 (the dashed line, open dot) there is
  directional selection for smaller \textsc{lma}.  At \textsc{lma} of
  0.08 (solid line, filled dot) there is a branching point (see inset
  for more detail).  (b) Landscape generated by two species, holding
  the first at \textsc{lma} of 0.08.  Introducing a species at the
  maximum fitness value from panel (a) reduces the region of positive
  fitness considerably (dashed line, open dot).  Dotted lines show
  subsequent invasions and replacements of the second species, the
  solid line shows a stable fitness landscape where both species exist
  on a peak in the fitness landscape.  At this point no further
  invasion is possible.  See Appendix \ref{sec:fitness} for further
  details and code.}
\label{fig:fitness}
\end{figure}

\clearpage
\setcounter{secnumdepth}{1}

\begin{appendices}\label{sec:appendices}

\section{Plant physiological model}\label{sec:FFW16}

See attached file \url{vignettes/physiology.pdf}

\section{Modelling demography of plants, patches and metapopulations}\label{sec:demography}

See attached file \url{vignettes/demography.pdf}

\section{Plant level properties}\label{sec:plant}

\emph{Note}: due to Manuscript Central not accepting \textsc{html}
files, we have had to convert this (and the following appendices) to
\textsc{pdf}s.  \textsc{html} versions are available on the wiki:
https://github.com/traitecoevo/plant\_paper/wiki --- the \textsc{html}
versions have less pixelated graphics and are easier to copy and
paste out of.

See attached file \url{vignettes/plant.pdf}, 
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/plant.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/plant}

\section{Cohort spacing algorithm}\label{sec:cohort-spacing}

See attached file \url{vignettes/cohort_spacing.pdf},
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/cohort_spacing.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/cohort_spacing}

\section{Finding demographic equilibrium}\label{sec:equilibrium}

See attached file \url{vignettes/equilibrium.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/equilibrium.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/equilibrium}

\section{Patch level dynamics}\label{sec:patch}

See attached file \url{vignettes/patch.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/patch.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/patch}

\section{Patch level emergent properties}\label{sec:emergent}

See attached file \url{vignettes/emergent.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/emergent.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/emergent}

\section{Calculating fitness}\label{sec:fitness}

See attached file \url{vignettes/fitness.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/fitness.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/fitness}

\section{Modifying parameters of the physiological model}\label{sec:parameters}
See attached file \url{vignettes/parameters.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/parameters.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/parameters}

\end{appendices}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-PDF-mode: t
%%% End:

%  LocalWords:  Westoby photosynthate Givnish Enquist Ruger feedbacks
%  LocalWords:  ontogenetic Shugart Tilman Geritz Hubbell Calcagno BH
%  LocalWords:  Cramer Sitch Dekauwe focussed successional Haxeltine
%  LocalWords:  Moorcroft lpj FFW lma wplcp Baltzer Lusk Kohyama pde
%  LocalWords:  Deroos Pacala Strigul ebt adaptively Coomes Binkley
%  LocalWords:  Ogawa Bormann Gyllenberg Vonfoerster Metz Dieckmann
%  LocalWords:  Rcpp Eddelbuettel templated RcppR testhat Wickham DS
%  LocalWords:  Schaling RG Brn nnstr Camac De Roos Duursma Johansson
%  LocalWords:  knitr